Parallel air palletization and tour planning for simultaneous pickup and delivery

### 2023-02-13 latex/comp_or.tex
Experimentation finished. Writing paper.

\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsthm}

@@

Once the assumptions of this work and the mathematical modelling of the problem are presented, it is easy to see that ACLP+RPDP is NP-hard. In a similar way to \cite[p. 6]{LurkinSchyns2015}, consider the simple case where $K=1$\/ (one leg), $m=2$ (two pallets around the aircraft CG), $2n$\/ sufficiently light items with same scores in $l_0$, and no items in $l_1$. Under these conditions, through polynomial reductions for the {\it Set-Partition Problem}, it is possible to demonstrate that the decision problem associated with ACLP+RPDP is NP-complete.

Real cases are more complex as they have hundreds of different items in each node and involve three intractable sub-problems: APP, WBP and SPDP. Through the mathematical modelling presented in the previous section, we verify that {\it Mixed-Integer Programming}\/ (MIP) is not able to solve these cases in feasible time. Thus, it is necessary to adopt some strategy to find a viable solution, not necessarily optimal, that seeks to maximize the objective function $f$.

@@

@@

		\Procedure{$ACLP+RPDP$}{$scenario,\ surplus,\ numProcs,\ timeLim$}

@@

			\For {each $\pi \in S_K$} \label{main:loop1}
				\State $f_{\pi} \gets SolveTour(\pi, L, M, C, N, method, numProcs,\ timeLim )$ \label{main:method}
			\EndFor \label{main:loop2}
			\State $answer[scenario,surplus,method] \gets \max f$ \label{main:f}
		\EndFor

@@

The other parameter $volume$\/ is a value greater than 1, which corresponds, at each node $l_k$, to the ratio between the sum of the volumes of the items ($\sum_{j=1}^{n_k} t^k_j.v$) and the load capacity of the pallets ($\sum_{i=1}^{m} p_i.v$). This parameter is passed to $ItemsGeneration$ (line \ref{main:items}), responsible for creating the items to be shipped, which will be presented in the next section.

@@

As we mentioned in the previous section, all tours start and end at $l_0$\/ (lines \ref{tour:pi1}-\ref{tour:pi2}).
After initializing the score and cost values (lines \ref{tour:score}-\ref{tour:cost}), there is a loop for the $K+1$\/ flights (lines \ref{tour:loop1}-\ref{tour:loop2}). Initially we set pallets destination as $-1$\/ (line \ref{tour:-1}). When the aircraft is at node $l_0$, the initial graph $G_1$\/ is empty because it has no consolidated items \ref{tour:g11}. Otherwise, the set $L_k$\/ of remaining nodes is updated (line \ref{tour:lk1}), and $UpdateConsolidated$\/ (line \ref{tour:dest}) returns the set of consolidated items that have not yet reached their destination and remain on board, rearranging them on the pallets to minimize CG deviation. This allocation is stored in graph $G_1$\/ (line \ref{tour:g12}).

In the context of this work, we know that $m>K$, once the aircraft has $7$\/ or $18$\/ pallets and $K\leq 6$, allowing there to be at least one pallet for each node to be visited. $SetPalletsDestination$\/ (line \ref{tour:dest2}) presets the destination of each pallet based on the volume demands of the current node, without changing the pallets destination with consolidated items.

@@

\begin{algorithm}[H]
	\caption{ Solves the sequence of nodes in the tour indexed by $\pi$ }  \label{alg:tour}

@@

		\Procedure{$SolveTour$}{$\pi, L, M, C, N, method, numProcs,\ timeLim$}

		\State $\pi(0) \gets 0$ \label{tour:pi1}
		\State $\pi(K+1) \gets 0$ \label{tour:pi2}
		\State $score \gets 0$ \label{tour:score}
		\State $cost \gets 0$ \label{tour:cost}
		\For {$k\gets0$ to $K$} \label{tour:loop1}		
			\For{$i \gets 1$ to $m$}
				\State $p_i.to[k] \gets -1$ \label{tour:-1}
			\EndFor	
			\If {$k = 0$}
				\State $L_0 \gets L$
				\State Let $G_1(M \cup N_0, \varnothing)$ \label{tour:g11}
			\Else
				\State $L_k \gets L_k - \pi(k)$  \label{tour:lk1}			
				\State $Q_{\pi(k)}, E^Q_{\pi(k)}, M \gets UpdateConsolidated(\pi(k))$ \label{tour:dest}			
				\State Let $G_1(M \cup N_{\pi(k)} \cup Q_{\pi(k)}, E^Q_{\pi(k)})$ \label{tour:g12}
			\EndIf  \label{tour:lk2}	
			\State $M \gets SetPalletsDestination( \pi(k) )$ \label{tour:dest2}		
			\State $G_2 \gets SolveNode(method, \pi(k), G_1, numProcs,\ timeLim)$ \label{tour:node}
			\State $s, \epsilon \gets ScoreAndDeviation(\pi(k), G_2)$ \label{tour:analyse}
			\State $score \gets score + s$ \label{tour:score2}
			\State $cost \gets cost + c_{\pi(k),\pi(k+1)} * (1 + c_g * |\epsilon|)$ \label{tour:cost2} 
		\EndFor  \label{tour:loop2}
		\Return $score / cost$ \label{tour:f}

@@

\subsection{Node-by-node solutions}
\label{methods}

In this subsection we present implementations of $SolveNode(...)$, where $method$\/ corresponds to a heuristic, $k$\/ is the index of the current node and $G$\/ is the allocation graph of the consolidated items that remain on board at node $l_k$.

@@

The selection of edges for $E^N_k$\/ uses the {\it edge attractiveness}\/ $\theta^k_{ij}$, Equation \ref{eq:edge}, which can be understood as the tendency to allocate the item $t^k_j$\/ to the pallet $p_i$. It is directly proportional to the score, and inversely proportional to the volume and the torque of each item.

\begin{equation} \label{eq:edge}
	\theta^k_{ij}= \frac{t^k_j.s}{t^k_j.v \times 3000}\times(1-\frac{t^k_j.w\times|p_i.d|}{\max_s\{t^k_s.w\}\times\max_q\{|p_q.d|\}});\ i \in \{1,2,\ldots,m\},\ j \in \{1,2,\ldots,n_k\}
\end{equation}

@@

In the pseudo-codes presented below, given an allocation graph $G$\/ for the node $l_k$, $f_s(G)$\/ is equivalent to the score $s$\/ returned by $ScoreAndDeviation(k,G) $, described in \cite{MesquitaSanches2023}.

@@

We also developed Algorithm \ref{alg:greedy}, which generates a greedy allocation of the items available in node $l_k$, according to the non-ascending order of $\theta^k_{ij}$, and considering the consolidated items already shipped ($E^Q_k$). Edges are included in this allocation as long as they respect feasibility constraints. Furthermore, the volume of each pallet $p_i$\/ cannot exceed $p_i.v \times limit$, where $ 0 < limit \leq 1$\/ is a given parameter.

		\Procedure{$Greedy$}{$k, G, limit$}

		\State Let $G(V_k, E^Q_k)$

		\State $E^N_k \gets \varnothing$

		\For{$q \gets 1$ to $m$}
			\If{$(p_i, a_q^k) \in E^Q_k$} 
				\State $volume[q] \gets volume[q] + a_q^k.v$ 
			\EndIf		
		\EndFor

		\For {each $e_{ij}^k$ in non-ascending order of $\theta^k_{ij}$}
			\If{($E^N_k \cup \{e_{ij}^k\}$ is feasible) {\bf and} ($volume[i] \leq p_i.v \times limit$)} 
				\State $E^N_k \gets E^N_k \cup \{e_{ij}^k\}$ 
				\State $volume[i] \gets volume[i] + t_j^k.v$
				\State $\eta_1[i] \gets \eta_1[i] + 1$ 	
			\EndIf
		\EndFor

		\Return $G(V_k, E^N_k \cup E^Q_k),\ \eta_1$ \Comment{return the solution and the last item indexes array }

\subsubsection{Multi-process Shims - mpShims}

@@

		\begin{tikzpicture}[scale=0.75, samples=100]
			\filldraw[fill=blue!3!white, draw=black] (0, 0) rectangle (12.5, 5.5);
			\draw[->] (.2, .6) -- coordinate (x axis mid) (12, .6);
			\node at (5, 0.3) {$\eta_1[i]$};
			\node at (5, 2.5) {|};
			\node at (6, 2.5) {shims};
			\node at (7, 2.5) {|};
			\node at (7, 0.3) {$\eta_2[i]$};
			\node at (0.3, 5) {$\theta_{ij}^k$};
			\draw[->] (.6, .2) -- coordinate (y axis mid) (0.6, 5.3);
			\node at (12, 0.3) {$e_{ij}^k$};
			\draw[smooth, domain = 0.09:2, color=black, thick] plot (.3+1/\x,{4.2+log2(\x)});
			\draw[->, dashed] (5, 0.6)--(5, 2.0);
			\draw[->, dashed] (7, 0.6)--(7, 1.5);
		\end{tikzpicture}

		\captionof{figure}{$n_k$\/ edges $e_{ij}^k$\/ of $p_i$\/ sorted by $\theta_{ij}^k$\/ in non-ascending order}
		\label{fig:whip}		
	\end{minipage}
\end{table}

Figure \ref{fig:whip} represents the $n_k$\/ possible edges $e_{ij}^k$\/ of $p_i$\/ sorted by $\theta_{ij}^k$. \emph{mpShims}\/ starts with a greedy solution, stopping at the edge with index in $\eta_1[i]$ close to the local optimum (first phase). Then, considering even the edge with the index $\eta_2[i]$, it elaborates different possible complements for this pallet (second phase) and selects the best of these complements (third phase).

\begin{algorithm}[H]
	\caption{Solve a node with {\it mpShims}}  \label{alg:mpshims}
	\begin{algorithmic}[1]

		\Procedure{$SolveNode$}{$mpShims, k, G, numProcs,\ timeLim$}

		\State $G^* \gets G(V_k, E^Q_k)$ \label{best_so_far}

		\State Sort $M$ by $|p_i.d|$ in non-descending order \label{mpshims:pallets}

		\State $E^N_k \gets \{\}$ \label{mpshims:items}

		\State Let $E[1..m][1..n_k]$, where $E[i]$ is an array with $n_k$ edges $(p_i,t^k_j)$ sorted by $\theta_{ij}^k$ in non-ascending order, $1 \leq i \leq m$ \label{mpshims:edges}

		\State $limits \gets getLimits(0.75,\ 0.98,\ numProcs)$ \label{mpshims:limits}

		\State $shimsQueue \gets \{\}$ \label{mpshims:queue}

		\For{$lim \in limits$}
			\State $shimsSolve(G(V_k, E^Q_k),\ M,\ shimsQueue,\ lim,\ E)$ \label{mpshims:send}
		\EndFor

		\State $counter \gets |limits|$

		\While{$counter > 0$}

			\State $G \gets \varnothing$

			\While{$runtime < timeLim$} \Comment{Wait until there is a solution in the queue}

			\State $G \gets shimsQueue.Pop()$
				\If{$G \ne \varnothing$}
					\Break
				\EndIf
			\EndWhile

			\If{$f_s(G) > f_s(G^*)$}
				\State $G^* \gets G$ 
			\EndIf
			\State $counter \gets counter - 1$

		\EndWhile

		\Return $G^*(V_k, E^N_k \cup E^Q_k)$ \label{mpshims:return}

Algorithm \ref{alg:mpshims} describes the {\it SolveNode} method when it receives {\it mpShims} as argument.

Line \ref{best_so_far} defines the best solution so far which is formed by a semi-loaded aircraft where pallets have been unloaded and some remain on board ($E^Q_k$) with consolidated items.

The Shims solution starts with the set of pallets closer to the center of gravity (\ref{mpshims:pallets}).

A new empty set of edges is created in line \ref{mpshims:items}.

Edges with items in node "k" to be loaded into not full pallets are defined in line \ref{mpshims:edges}.

Line \ref{mpshims:limits} creates a list of unique limits with values between 0.75 and 0.98, that will be sent to different parallel processes to generate diverse solutions.

A queue of shims solutions to be used by the child processes to send solution to the main process is defined in line \ref{mpshims:queue}.

The iteration in \ref{mpshims:send} publishes or triggers parallel processes to solve $G(V_k, E^Q_k)$.

In line \ref{mpshims:wait} the program hangs until each child process put a solution into the queue. Each solution is collected by the main process and compared with the best so far.

Finally, in line \ref{mpshims:return}, the best solution found is returned.

\begin{algorithm}[H]
	\caption{Shims solving method to be run in parallel}  \label{alg:shimsSolve}
	\begin{algorithmic}[1]

		\Procedure{$shimsSolve$}{$G,\ M,\ shimsQueue,\ limit,\ E$}

		\State $G(V_k,\ E^Q_k \cup E^N_k),\ \eta_1 \gets Greedy(k,\ G(V_k,\ E^Q_k),\ limit)$ \label{first_phase}

		\For{$i \gets 1$ to $m$}

			\State $\eta_2[i] \gets \eta_1[i]$

			\Repeat \label{second_phase_begin}
				\State $\eta_2[i] \gets \eta_2[i] + 1$ 
				\State $e^k_{ij} \gets E[i][\eta_2[i]]$
				\State $volume \gets volume + t^k_j.v$
			\Until{($\eta_2[i] = n_k$) {\bf or} ($ volume \geq (2 - limit)*p_i.v)$)} \label{second_phase_end}

			\State $slack[i] \gets (1-limit) \times p_i.v$

			\State $E^N_k \gets E^N_k \cup getBestShims(i,\ \eta_1,\ \eta_2,\ E,\ k,\ slack[i])$ \label{third_phase}

@@

		\State $shimsQueue \gets shimsQueue \cup \{G(V_k, E^N_k \cup E^Q_k)\}$ \label{enqueue}

@@

Algorithm \ref{alg:shimsSolve}, line \ref{first_phase} corresponds to Shims' first phase. Lines  \ref{second_phase_begin}-\ref{second_phase_end} correspond to the Shims' second phase.
Line \ref{third_phase} corresponds to Shims' third phase. In the last line, \ref{enqueue}, {\it mpShims} puts the solution into the queue to be retrieved by the main process.

\begin{algorithm}[H]
	\caption{Get the best shims of edges that fills each pallet gap}  \label{alg:getBestShims}
	\begin{algorithmic}[1]

@@

\multirow{7}{*}{$1.2$}&\multirow{2}{*}{ $Gurobi$}& $f$                & 12.00       & {\bf 8.38}  & {\bf 12.19} & 13.38       &  14.42      & 13.25        & 0.96 \\% 7651
&                                                & {\it run-time (s)} &  8          &   14        &  25         &  95         &    217      &  1719        & 1.0  \\%2078

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $Shims$}                       & $f$                & 8.37        &  6.40       &  7.68       &  8.05       &   7.72      & 6.74         & 0.56 \\%
&                                                & {\it run-time (s)} &  < 1        &  < 1        &    3        &   11        &   27        &  342         & 5.4   \\%385

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $mpShims$}                     & $f$                & {\bf 12.24} &  7.95       &  11.59      & {\bf 13.80} & {\bf 16.16} & {\bf 13.74}  & {\bf 0.99} \\%
&                                                & {\it run-time (s)} &  < 1        &  < 1        &    2        &   6         &    15       &   165        & {\bf 10.9}  \\% 190

@@

\multirow{7}{*}{$1.5$}&\multirow{2}{*}{ $Gurobi$}& $f$                & 16.49       & {\bf 11.60} & {\bf 16.68} & 18.33       &  20.10      & 17.94        & 0.97  \\%10402
&                                                & {\it run-time (s)} &  12         &   25        &  45         &  169        &   437       &  2377        & 1.0  \\%3065

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $Shims$}                       & $f$                & 12.53       &  8.31       &  10.04      &  10.94      &   11.72     & 10.13        & 0.61 \\%
&                                                & {\it run-time (s)} &  < 1        &  3          &    6        &   24        &   70        &  419         & 5.9 \\%523

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $mpShims$}                     & $f$                & {\bf 16.71} & 10.86       & {\bf 16.68} & {\bf 18.84} & {\bf 21.30} & {\bf 19.39}  & {\bf 1.00}  \\%
&                                                & {\it run-time (s)} &  < 1        &  2          &    3        &   9         &   26        &   158        & {\bf 15.4}  \\%199

\midrule
\multirow{7}{*}{$2.0$}&\multirow{2}{*}{ $Gurobi$}& $f$                & 24.04       & {\bf 17.33} & 24.62       & {\bf 26.36} &  28.88      & 25.89        & 0.98  \\%14962
&                                                & {\it run-time (s)} &  14         &   34        &  62         &  236        &   609       & 3431         & 1.0  \\%4386

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $Shims$}                       & $f$                & 18.41       &  13.37       &  15.90      &  16.17      &   17.08     & 15.96        & 0.65 \\%
&                                                & {\it run-time (s)} &  < 1        &  4          &    7        &   30         &   85        &  501         & 7.0 \\%628

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $mpShims$}                     & $f$                & {\bf 24.28} & 16.54       & {\bf 25.53} & 25.87        & {\bf 29.10} & {\bf 27.02}  & {\bf 0.99} \\%
&                                                & {\it run-time (s)} &  < 1        &  2          &    3        &   12         &   32        &   188        & {\bf 18.4}  \\%238

@@

For the MIP solver in some scenarios, less than 50\% of the results were optimal, and the remaining were suboptimal due to the time limit. This explains that {\it mpShims} surpassed the MIP solver results.

>>>>

\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{amsmath}

@@

Once the assumptions of this work and the mathematical modelling of the problem are presented, it is easy to see that ACLP+RPDP is NP-hard. In a similar way to \cite[p. 6]{LurkinSchyns2015}, consider the simple case where $K=1$\/ (one leg), $m=2$ (two pallets around the aircraft CG), $2n$\/ sufficiently light items with same scores in $l_0$, and no items in $l_1$. Under these conditions, through polynomial reductions for the {\it Set-Partition Problem}, it is possible to demonstrate that the decision problem associated with ACLP+RPDP is NP-hard.

Real cases are more complex as they have hundreds of different items in each node and involve three intractable sub-problems: APP, WBP and SPDP. Through the mathematical modeling presented in the previous section, we verify that {\it Mixed-Integer Programming}\/ (MIP) is not able to solve these cases in feasible time. Thus, it is necessary to adopt some strategy to find a viable solution, not necessarily optimal, that seeks to maximize the objective function $f$.

@@

{\color{blue} Each node, except the base, inherits information from the previous nodes (pallets that must remain on board), and must solved taking into account the remaining not visited nodes. One more reason for a node-by-node approach.}

@@

		\Procedure{$ACLP+RPDP$}{$scenario,\ surplus,\ timeLim$}

@@

			\For {each $\pi \in S_K$} \label{main:loop1}
				\State $f_{\pi} \gets SolveTour(\pi, L, M, C, N, method, \ timeLim )$ \label{main:method}
			\EndFor \label{main:loop2}
			\State $answer[scenario,surplus,method] \gets \max f$ \label{main:f}
		\EndFor

@@

The other parameter $volume$\/ is a value greater than 1, which corresponds, at each node $l_{\pi(k)}$, to the ratio between the sum of the volumes of the items ($\sum_{j=1}^{n_{\pi(k)}} t^{\pi(k)}_j.v$) and the load capacity of the pallets ($\sum_{i=1}^{m} p_i.v$). This parameter is passed to $ItemsGeneration$ (line \ref{main:items}), responsible for creating the items to be shipped, which will be presented in the next section.

@@

As we mentioned in the previous section, all tours start and end at $l_0$\/ (lines \ref{tour:pi1}-\ref{tour:pi2}).
After initializing the score and cost values (lines \ref{tour:score}-\ref{tour:cost}), there is a loop for the $K+1$\/ flights (lines \ref{tour:loop1}-\ref{tour:loop2}). Initially we set pallets destination as $-1$\/ (line \ref{tour:-1}). When the aircraft is at node $l_0$, the initial graph $G_1$\/ is empty because it has no consolidated items \ref{tour:g11}. Otherwise, the set $L_k$\/ of remaining nodes when departing from $\pi(k)$ is updated (line \ref{tour:lk1}), and $UpdateConsolidated$\/ (line \ref{tour:dest}) returns the set of consolidated items that have not yet reached their destinations and remain on board, rearranging them on the pallets to minimize CG deviation. This allocation is stored in graph $G_1$\/ (line \ref{tour:g12}).

In the context of this work, we know that $m>K$, once the aircraft has $7$\/ or $18$\/ pallets and $K\leq 6$, allowing there to be at least one pallet for each node to be visited. $SetPalletsDestination$\/ (line \ref{tour:dest2}) presets the destination of each pallet based on the volume demands caused by the items to be embarked in the current node, without changing the pallets destination with consolidated items.

@@

\begin{algorithm}[H]
	\caption{ Solves the sequence of nodes of tour $\pi$ }  \label{alg:tour}

@@

		\Procedure{$SolveTour$}{$\pi, L, M, C, N, method, timeLim$}

		\State $\pi(0) \gets 0$ \label{tour:pi1}
		\State $\pi(K+1) \gets 0$ \label{tour:pi2}
		\State $score \gets 0$ \label{tour:score}
		\State $cost \gets 0$ \label{tour:cost}
		\For {$k\gets0$ to $K$} \label{tour:loop1}		
			\For{$i \gets 1$ to $m$}
				\State $p_i.to[k] \gets -1$ \label{tour:-1} \Comment{reset this pallet destination}
			\EndFor	
			\If {$k = 0$}
				\State $L_0 \gets L$
				\State Let $G_1(M \cup N_0, \varnothing)$ \label{tour:g11}
			\Else
				\State $L_k \gets L_k - \pi(k)$  \label{tour:lk1}			
				\State $Q_{\pi(k)}, E^Q_{\pi(k)}, M \gets UpdateConsolidated(\pi(k))$ \label{tour:dest}			
				\State Let $G_1(M \cup N_{\pi(k)} \cup Q_{\pi(k)}, E^Q_{\pi(k)})$ \label{tour:g12}
			\EndIf  \label{tour:lk2}	
			\State $M \gets SetPalletsDestination( \pi(k) )$ \label{tour:dest2}		
			\State $G_2 \gets SolveNode(method, \pi(k), G_1, timeLim)$ \label{tour:node}
			\State $s, \epsilon \gets ScoreAndDeviation(\pi(k), G_2)$ \label{tour:analyse}
			\State $score \gets score + s$ \label{tour:score2}
			\State $cost \gets cost + c_{\pi(k),\pi(k+1)} * (1 + c_g * |\epsilon|)$ \label{tour:cost2} 
		\EndFor  \label{tour:loop2}
		\Return $score / cost$ \label{tour:f}

@@

\subsection{Node-by-node resolution}
\label{methods}

In this subsection we present implementations of $SolveNode(...)$, where $method$\/ corresponds to a heuristic, $k$\/ is the index of the current node $l_{\pi(k)}$ and $G$\/ is the allocation graph of the consolidated items that remain on board at $l_{\pi(k)}$.

@@

The selection of edges for $E^N_{\pi(k)}$\/ uses the {\it edge attractiveness}\/ $\theta^{\pi(k)}_{ij}$, Equation \ref{eq:edge}, which can be understood as the tendency to allocate the item $t^{\pi(k)}_j$\/ to the pallet $p_i$. It is directly proportional to the score, and inversely proportional to the volume and the torque of each item.

\begin{equation} \label{eq:edge}
	\theta^{\pi(k)}_{ij}= \frac{t^{\pi(k)}_j.s}{t^{\pi(k)}_j.v \times 3000}\times(1-\frac{t^{\pi(k)}_j.w\times|p_i.d|}{\max_s\{t^{\pi(k)}_s.w\}\times\max_q\{|p_q.d|\}});\ i \in \{1,2,\ldots,m\},\ j \in \{1,2,\ldots,n_k\}
\end{equation}

@@

In the pseudo-codes presented below, given an allocation graph $G$\/ for the node $l_{\pi(k)}$, $f_s(G)$\/ is equivalent to the score $s$\/ returned by $ScoreAndDeviation({\pi(k)},G) $, described in \cite{MesquitaSanches2023}.

@@

\subsection{Multi-process Shims - mpShims}

@@

		\begin{tikzpicture}[scale=0.75, samples=100]
			\filldraw[fill=blue!3!white, draw=black] (0, 0) rectangle (12.5, 5.5);
			\draw[->] (.2, .6) -- coordinate (x axis mid) (12, .6);
			\node at (5, 0.3) {$\eta_1[i]$};
			\node at (5, 2.5) {|};
			\node at (6, 2.5) {shims};
			\node at (7, 2.5) {|};
			\node at (7, 0.3) {$\eta_2[i]$};
			\node at (0.3, 5) {$\theta_{ij}^{\pi(k)}$};
			\draw[->] (.6, .2) -- coordinate (y axis mid) (0.6, 5.3);
			\node at (12, 0.3) {$e_{ij}^{\pi(k)}$};
			\draw[smooth, domain = 0.09:2, color=black, thick] plot (.3+1/\x,{4.2+log2(\x)});
			\draw[->, dashed] (5, 0.6)--(5, 2.0);
			\draw[->, dashed] (7, 0.6)--(7, 1.5);
		\end{tikzpicture}

		\captionof{figure}{$n_{\pi(k)}$\/ edges $e_{ij}^{\pi(k)}$\/ of $p_i$\/ sorted by $\theta_{ij}^{\pi(k)}$\/ in non-ascending order}
		\label{fig:whip}		
	\end{minipage}
\end{table}

Figure \ref{fig:whip} represents the $n_{\pi(k)}$\/ possible edges $e_{ij}^{\pi(k)}$\/ of $p_i$\/ sorted by $\theta_{ij}^{\pi(k)}$. \emph{Shims}\/ starts with a greedy solution, stopping at the edge with index in $\eta_1[i]$ close to the local optimum (first phase). Then, considering even the edge with the index $\eta_2[i]$, it elaborates different possible complements for this pallet (second phase) and selects the best of these complements (third phase).

{\color{blue} In this work, we included a new phase between the first and the second: a minimization of the CG deviation. An integer programming procedure that helps improve the overall {\it mpShims} results.

$1^{st}$ - Greedy phase

$2^{st}$ - Minimization of CG deviation

$3^{st}$ - Shims construction

$4^{st}$ - best Shims selection

All executed in a multi-process runtime, accessing concurrently, reading and writing, the list of items and the current value of the CG deviation.

As to the multi-process runtime, we considered each pallet as a process.
}

\subsection{mpShims}

\begin{algorithm}[H]
	\caption{Mount a greedy solution until the volume limit for each pallet}  \label{alg:mpShims}

		\Procedure{$mpShims$}{${N_{\pi(k)}, \pi(k)}, limit, G$}

		\State Let $G(E^{Q_{\pi(k)}})$ \Comment{the solution comes with some consolidated on board}

		\State $surplus \gets 1 + 3 \times (1 - limit)$

		\State $nodeTorque \gets 0$

		\State $ts \gets 2.0$ \Comment{torque surplus} \label{mpShims:ts}

		\State $lock \gets$ multi-processing lock \Comment{for shared data among processes to avoid race condition}

		\State $proc \gets \{p_1,p_2,..., p_m\}$  \Comment{each pallet has its own process}

		\For{$i \gets 1$ to $m$}
		\State $proc_i \gets Greedy(p_i,\ N_{\pi(k)},\ \pi(k),\ nodeTorque,\ G,\ limit,\ lock,\ ts )$
		\State $proc_i.start$
		\EndFor

		\For{$i \gets 1$ to $m$}
		\State $proc_i.join$
		\EndFor

		\State $minCGdev(M,\ \pi(k),\ nodeTorque)$  \label{mpShims:minCGdev} \Comment{minimize the CG deviation.}

		\For{$i \gets 1$ to $m$}
		\State $proc_i \gets getBestShims(p_i,\ N_{\pi(k)},\ \pi(k),\ nodeTorque,\ G,\ surplus,\ lock,\ ts )$
		\State $proc_i.start$
		\EndFor

		\For{$i \gets 1$ to $m$}
		\State $proc_i.join$

@@

		\Return $G(E^{N_{\pi(k)}} \cup E^{Q_{\pi(k)}})$

@@

Line \ref{mpShims:ts} defines the torque surplus ($ts$). For this phase, the solution may generate torque infeasible solutions with doubled torque, which will be repaired by the next phase, the minimization of the CG deviation.

Line \ref{mpShims:minCGdev} presents the procedure to minimize the CG deviation and repair the infeasible greedy solution.

\subsubsection{Shims first phase}

Algorithm \ref{alg:greedy} generates a greedy allocation of the items available in node $l_{\pi(k)}$, according to the non-ascending order of $\theta^{\pi(k)}_{ij}$, and considering the consolidated items already shipped ($E^Q_{\pi(k)}$). Edges are included in this allocation as long as they respect feasibility constraints. Furthermore, the volume of each pallet $p_i$\/ cannot exceed $p_i.v \times limit$, where $ 0 < limit \leq 1$\/ is a given parameter.

		\Procedure{$Greedy$}{${\pi(k)}, G, limit$}

		\State Let $G(V_{\pi(k)}, E^Q_{\pi(k)})$

		\State $E^N_{\pi(k)} \gets \varnothing$

		\For{$q \gets 1$ to $m$}
		\If{$(p_i, a_q^{\pi(k)}) \in E^Q_{\pi(k)}$} 
		\State $volume[q] \gets volume[q] + a_q^{\pi(k)}.v$ 
		\EndIf		
		\EndFor

		\For {each $e_{ij}^{\pi(k)}$ in non-ascending order of $\theta^{\pi(k)}_{ij}$}
		\If{($E^N_{\pi(k)} \cup \{e_{ij}^{\pi(k)}\}$ is feasible) {\bf and} ($volume[i] \leq p_i.v \times limit$)} 
		\State $E^N_{\pi(k)} \gets E^N_{\pi(k)} \cup \{e_{ij}^{\pi(k)}\}$ 
		\State $volume[i] \gets volume[i] + t_j^{\pi(k)}.v$
		\State $\eta_1[i] \gets \eta_1[i] + 1$ 	
		\EndIf
		\EndFor

		\Return $G(V_{\pi(k)}, E^N_{\pi(k)} \cup E^Q_{\pi(k)}),\ \eta_1$ \Comment{return the partial solution and the last item indexes array }

\subsubsection{Shims second phase: improvement on minimizing CG deviation}

Between the greedy and the {\it Shims} selection phases we inserted, in this work, a method to minimize the CG deviation, as stated by the model that follows.

Let $Q^{\pi(k)}  = \{ a^{\pi(k)}_1, a^{\pi(k)}_2, \ldots, a^{\pi(k)}_m \}$ be the set of consolidated items assembled on the set of $m$ pallets in node $\pi(k)$. There is a bi-univocal relation between pallets and consolidated.

Let $H^{\pi(k)}_{iq}$ be the set of decision variables that relates pallet $i$ to consolidated $q$ in node $\pi(k)$.

\begin{equation} \label{eq:distobjshims}
	\mbox{minimize}\ \Bigg |  \sum_{i=1}^{m} \sum_{q=1}^{m} H^{\pi(k)}_{iq} \times  (140 + a^{\pi(k)}_q.w) * p_i.d\ \Bigg | 
\end{equation}

\begin{equation} \label{eq:distconsshims}
	s.t.: \sum_{i=1}^{m} H^{\pi(k)}_{iq} = 1;\ q \in \{1,2,\ldots,m\}
\end{equation}

\begin{equation} \label{eq:distpalletsshims}
	s.t.: \sum_{q=1}^{m} H^{\pi(k)}_{iq}=1;\ i \in \{1,2,\ldots,m\}
\end{equation}

Equation \ref{eq:distobjshims} minimizes the aircraft absolute torque in node $\pi(k)$, where 140 is the weight of an empty pallet, $a^{\pi(k)}_q.w$ the weight of the consolidated, and $p_i.d$ the distance between the pallet centroid to the aircraft CG; Equation \ref{eq:distconsshims} states that each consolidated will be assigned to exactly one pallet; and Equation \ref{eq:distpalletsshims} states that each pallet will receive one consolidated.

\subsubsection{Shims third and fourth phases}

\begin{algorithm}[H]
	\caption{Mount shims of edges that fills each pallet gap and return the best shims}  \label{alg:getBestShims}
	\begin{algorithmic}[1]

@@

\multirow{7}{*}{$1.2$}&\multirow{2}{*}{ $Gurobi$}& $f$                & {\bf 12.00} & {\bf 8.38}  & {\bf 12.19} & {\bf 13.38} & {\bf 14.42} & 13.25        & {\bf 0.99} \\% 7412
&                                                & {\it run-time (s)} &  8          &   14        &  25         &  95         &    217      &  1719        &  1.0       \\%2078

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $Shims$}                       & $f$                & 9.66        &  6.37       &  9.58       &  10.87      &   12.24     & 11.94        &  0.82\\%
&                                                & {\it run-time (s)} &  < 1        &  3          &    5        &   11        &   27        & 342          &  5.3\\%389

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $mpShims$}                     & $f$                & 11.90       &  7.61       &  10.86      & 13.13       & 13.75       & {\bf 13.75}  &  0.96\\%
&                                                & {\it run-time (s)} &  < 1        &  < 1        &    3        &  9          &  26         & 156          &  {\bf 10.6}\\%

@@

\multirow{7}{*}{$1.5$}&\multirow{2}{*}{ $Gurobi$}& $f$                & 16.49       & {\bf 11.60} & {\bf 16.68} & {\bf 18.33} & {\bf 20.10} & 17.94        & {\bf 0.99}\\%10209
&                                                & {\it run-time (s)} &  12         &   25        &  45         &  169        &   437       &  2377        & 1.0\\%3065

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $Shims$}                       & $f$                & 13.71       &  9.06       &  13.54      &  15.21      &   17.09     &  17.15       & 0.84\\%
&                                                & {\it run-time (s)} &  < 1        &  3          &    6        &   24        &   70        &  419         & 5.9\\%

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $mpShims$}                     & $f$                & {\bf 16.85} & 10.65       & {\bf 16.68} & 17.05       & 19.72       & {\bf 18.53}  & 0.98\\%
&                                                & {\it run-time (s)} &  < 1        &  2          &    3        &   10        &   28        &   173        & {\bf 14.1}\\%

\midrule
\multirow{7}{*}{$2.0$}&\multirow{2}{*}{ $Gurobi$}& $f$                & {\bf 24.04} & {\bf 17.33} & {\bf 24.62} & {\bf 26.36} & {\bf 28.88} & {\bf 25.89}  & {\bf 1.00}\\%14712
&                                                & {\it run-time (s)} &  14         &   34        &  62         &  236        &  609        & 3431         & 1.0\\%4836

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $Shims$}                       & $f$                & 20.42       &  13.89      &  20.36      &  22.71      &   25.13     & 24.69        & 0.86\\%
&                                                & {\it run-time (s)} &  < 1        &  4          &    7        &   30        &   85        &  501         & 7.7\\%

\cmidrule{2-10}		                       
&\multirow{2}{*}{ $mpShims$}                     & $f$                & 23.12       & 14.89       & 22.08       & 23.96       & 26.98       & 25.29        &  0.93\\%
&                                                & {\it run-time (s)} &  < 1        &  2          &    3        &   12        &   33        &  203         & {\bf 19.0}\\%

@@

For the MIP solver in some scenarios, less than 50\% of the results were optimal, and the remaining were suboptimal due to the time limit. This explains that {\it mpShims} surpassed some of the MIP solver results.

In \cite{MesquitaSanches2023}, {\it Shims} surpassed the MIP solver, on average, because the time limit they used (0.7s) was sufficient to guarantee a worst-case solution in less than 30min. This decision caused most of the MIP solver results to be poor, making {\it Shims} outstanding as it was the unique method to perform well with all scenarios and instances.
